---
title: "Reproducible Research - Peer Assessment I"
author: "Ernesto Aguirre"
date: "Saturday, June 13, 2015"
output: html_document
---

Loading and preprocessing the data

Show any code that is needed to

  1.-  Load the data (i.e. read.csv())

  2.-  Process/transform the data (if necessary) into a format suitable for your analysis


```{r}
setwd("C:/R-Neto/Asg1-51/data")
activity <- read.csv("activity.csv", colClasses = c("numeric", "character","numeric"))
head(activity)
names(activity)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```
Open de lattice library (We will need it later)

```{r}
library(lattice)
```
What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

    Calculate the total number of steps taken per day

    Calculate and report the mean and median of the total number of steps taken per day


```{r, echo=TRUE}
StepsTotal <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
hist(StepsTotal$steps, main = "Total steps by day", xlab = "day", col = "red")
```

Mean
```{r, echo=TRUE}
mean(StepsTotal$steps)
```

Median
```{r, echo=TRUE}
median(StepsTotal$steps)
```

What is the average daily activity pattern?

    Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

    Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = T)
length(stepsday)
for (i in 1:61) {
steps[i] <- stepsday[[i]]
day[i] <- names(stepsday)[i]
}
df <- data.frame(day, steps)
head(df)
hist(df$steps, main = "Total steps by day", xlab = "day", col = "green")
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", ylab = "Average across all Days", main = "Average number of steps taken", col = "red")
max_interval <- which.max(time_series)
names(max_interval)
```

Imputing missing values

    Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

    Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

    Create a new dataset that is equal to the original dataset but with the missing data filled in.

    Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r, echo=TRUE}
activity_NA <- sum(is.na(activity))
activity_NA
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
obs <- activity[i, ]
if (is.na(obs$steps)) {
steps <- subset(StepsAverage, interval == obs$interval)$steps
} else {
steps <- obs$steps
}
fillNA <- c(fillNA, steps)
}
new_activity <- activity
new_activity$steps <- fillNA
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")
```

Mean
```{r, echo=TRUE}
mean(StepsTotal2$steps)
```

Median
```{r, echo=TRUE}
median(StepsTotal2$steps)
```

Are there differences in activity patterns between weekdays and weekends?

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
sum(is.na(activity$steps))


```
Create a new dataset with the missing data filled in.



```{r, echo=TRUE}
steps.interval <- aggregate(steps ~ interval, data = activity, FUN = mean)
steps.interval$interval[which.max(steps.interval$steps)]
activity <- merge(activity, steps.interval, by = "interval", suffixes = c("", ".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity <- activity[, c(1:3)]
steps.date <- aggregate(steps ~ date, data = activity, FUN = sum)
## barplot(steps.date$steps, names.arg = steps.date$date, xlab = "date", ylab = "steps")
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
activity$daytype <- as.factor(sapply(activity$date, daytype))
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    steps.type <- aggregate(steps ~ interval, data = activity, FUN = mean)
    
    plot(steps.type, type = "l", main = type)
}

```